---
title: "Escapement Goals Evaluation"
author: "Ibrahim Ilhan"
date: "10/17/2018"
output: html_document
---

```{r setup, include=FALSE}
# setup, include=FALSE
knitr::opts_chunk$set(echo = TRUE)
```

# Preparing Daily Escapment Data
Alaska Department of Fish and Game. 2017. Daily salmon escapement counts (ADFG_firstAttempt_reformatted.csv) from the OceanAK database, Alaska, 1921-2017. Knowledge Network for Biocomplexity. doi:10.5063/F1TX3CKH.

+ Reading data from url
```{r}
daily_escapement_data <- read.csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Af119a05b-bbe7-4aea-93c6-85434dcb1c5e")
```

+ Viewing what's in the file (first six rows)
```{r}
head(daily_escapement_data)
#tail(daily_escapement_data)
```
+ Necessary libraries for packages that will be used
```{r warning=FALSE, message=FALSE}
library(dplyr)
library(tidyr)
```

# Preparing Sockeye Escapment Goals
+ Printing unique information within the SASAP.Region
```{r}
unique(daily_escapement_data$SASAP.Region)
```

+ Filtering region (Bristol Bay) and type of salmon (Sockeye) from daily_escapement_data
```{r}
daily_escapement_sockeye <- daily_escapement_data %>% 
  filter(Species == "Sockeye") %>% 
  filter(SASAP.Region == "Bristol Bay")
  head(daily_escapement_sockeye)
```
# Evaluating Sockeye Escapment Goals
+ Summarizing yearly counts of Sockeye for each river
```{r}
daily_escapement_sockeye_sum <- daily_escapement_sockeye %>% 
  separate(sampleDate, into = c("year", "month", "day"), sep = "-") %>%
  group_by(Location, year) %>%
  summarise(annual_count = sum(DailyCount))

  head(daily_escapement_sockeye_sum)
```

```{r}
goals <- read.csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/knb.92014.1")

  head(goals)
```

```{r}
goals_clean <- goals %>%
  select(Species, Region, System, Lower, Upper, Type, Initial.Year) %>%  
  filter(Region == "Bristol Bay" & Species == "Sockeye") %>% 
  mutate(Lower = as.numeric(Lower), Initial.Year = as.numeric(Initial.Year))

head(goals)
```
```{r}
goals_joined <- left_join(esc_clean, goals_clean, by = c("Lcation" = "System")) %>% 
  #mutate(goal_met = ifelse(annual_count > Lower & annual_count < Upper, TRUE, FALSE)) %>% 
  #select(-Species, -Region)
head(goals_joined)
```

