---
title: "Data Tydying"
author: "Ibrahim Ilhan"
date: "10/16/2018"
output: html_document
---

#7.3 Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning = F, message = F}
library(dplyr)
library(tidyr)
```

# Read and catch data from: ... head(catch_df), lists first 6 rows of the read.csv; catch_df, is read from the url with a method of libcurl 

```{r}
catch_df <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", method = "libcurl"), stringsAsFactors = FALSE)
head(catch_df)
```

#7.4 About the pipe (%>%) operator | command+shift+m = %>% 
# If you think of the assignment operator (<-) as reading like “gets”, then the pipe operator would read like “then.”

'select' the columns we need for analysis

```{r}
catch_clean <- catch_df %>%
  #select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum)
  select(-All, -notesRegCode)

head(catch_clean)
```

```{r}
catch_clean <- catch_df %>%
  #select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum)
  select(-All, -notesRegCode) %>%      # -disselect those columns (e.g. All and noteRegCode)
  gather(key = "species", value = "catch", -Region, -Year)

head(catch_clean)
```

```{r}
catch_clean <- catch_df %>%
  #select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum)
  select(-All, -notesRegCode) %>%      # -disselect those columns (e.g. All and noteRegCode)
  gather(key = "species", value = "catch", -Region, -Year)

#head(catch_clean)
tail(catch_clean)
```

```{r}
catch_clean <- catch_df %>%
  #select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum)
  select(-All, -notesRegCode) %>%      # -disselect those columns (e.g. All and noteRegCode)
  gather(key = "species", value = "catch", -Region, -Year) %>% 
  rename(catch_thousands = catch)

#head(catch_clean)
tail(catch_clean)
```

```{r}
test <- as.numeric(catch_clean$catch_thousands)
i <- which(is.na(test) == T)

catch_clean[i, ]
```


```{r}
catch_clean_mutated <- catch_clean %>%
  mutate(catch_thousands = ifelse(catch_thousands == "I", 1, catch_thousands)) %>% 
  mutate(catch_thousands = as.numeric(catch_thousands)) %>% 
  mutate(catch = catch_thousands * 1000) %>%
  select(-catch_thousands)

tail(catch_clean_mutated)
```

```{r}
catch_clean_mutated %>%
  filter(species == "Chinook") %>% 
  arrange(-catch)
  #arrange(catch)
  #group_by(Region) %>%
  #summarise(mean(catch))
  #7.8-7.11
```

# 7.12 Joins in dplyr
# load region data table
```{r}
region_defs <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1", method = "libcurl"), stringsAsFactors = FALSE)

#read.ssv("user/home/...") | reading from a local directory

head(region_defs)
```

```{r}
catch_joined <- left_join(catch_clean_mutated, region_defs, by = c("Region" = "code"))

head(catch_joined)
#nrow(catch_joined)
#nrow(catch_clean_mutated)
```

```{r}
dates_df <- data.frame(date = c("5/24/1930",
                                "5/25/1930",
                                "5/26/1930",
                                "5/27/1930",
                                "5/28/1930"),
                       stringsAsFactors = FALSE)

dates_df
```

#7.13 separate() and unite()
```{r}
dates_separate <- dates_df %>%
  separate(date, into = c("month", "day", "year"), remove = F) %>% 
  unite(date_isl, year, month, day, sep = "-")

dates_separate
```

