---
title: "Data Tydying"
author: "Ibrahim Ilhan"
date: "10/16/2018"
output: html_document
---

#7.3 Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning = F, message = F}
library(dplyr)
library(tidyr)
```

# Read and catch data from: ... head(catch_df), lists first 6 rows of the read.csv; catch_df, is read from the url with a method of libcurl 

```{r}
catch_df <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", method = "libcurl"), stringsAsFactors = FALSE)
head(catch_df)
```

#7.4 About the pipe (%>%) operator | command+shift+m = %>% 
# If you think of the assignment operator (<-) as reading like “gets”, then the pipe operator would read like “then.”

'select' the columns we need for analysis

```{r}
catch_clean <- catch_df %>%
  #select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum)
  select(-All, -notesRegCode)

head(catch_clean)
```

```{r}
catch_clean <- catch_df %>%
  #select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum)
  select(-All, -notesRegCode) %>%      # -disselect those columns (e.g. All and noteRegCode)
  gather(key = "species", value = "catch", -Region, -Year)

head(catch_clean)
```

```{r}
catch_clean <- catch_df %>%
  #select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum)
  select(-All, -notesRegCode) %>%      # -disselect those columns (e.g. All and noteRegCode)
  gather(key = "species", value = "catch", -Region, -Year)

#head(catch_clean)
tail(catch_clean)
```
```{r}
catch_clean <- catch_df %>%
  #select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum)
  select(-All, -notesRegCode) %>%      # -disselect those columns (e.g. All and noteRegCode)
  gather(key = "species", value = "catch", -Region, -Year) %>% 
  rename(catch_thousands = catch)

#head(catch_clean)
tail(catch_clean)
```


